---
title: "VPD Surveillance in Yemen (2024–2025)"
output:
  flexdashboard::flex_dashboard:
    theme: flatly
    orientation: rows
    vertical_layout: scroll
    navbar:
      - { title: "GitHub Source Code", href: "https://github.com/Akanshya1998/DATA555_Dashboard", align: right, icon: "fa-github" }
    css: |
      /* === GLOBAL === */
      body {
        font-family: Arial, sans-serif;
        font-size: 15px;
        color: #2c3e50;
        background-color: #f5f6fa;
      }

      /* === NAVBAR === */
      .navbar, .navbar-default {
        background-color: #1a5276 !important;
        border: none !important;
      }
      .navbar-brand,
      .navbar-default .navbar-brand,
      .navbar-nav > li > a,
      .navbar-default .navbar-nav > li > a {
        color: #ffffff !important;
        font-weight: 700 !important;
        font-size: 15px !important;
      }
      .navbar-default .navbar-nav > .active > a,
      .navbar-default .navbar-nav > .active > a:hover {
        background-color: #154360 !important;
        color: #ffffff !important;
      }
      .navbar-nav > li > a:hover {
        background-color: #1f618d !important;
      }

      /* === VALUE BOXES === */
      .value-box .value {
        font-size: 30px !important;
        font-weight: 700 !important;
      }
      .value-box .caption {
        font-size: 12.5px !important;
        font-weight: 500 !important;
        line-height: 1.4 !important;
      }

      /* === CALLOUT BOX (how to read) === */
      .callout {
        background-color: #eaf4fb;
        border-left: 5px solid #2471A3;
        border-radius: 5px;
        padding: 14px 18px;
        font-size: 14px;
        line-height: 1.8;
        margin: 6px 0 10px 0;
        color: #1a2f45;
      }

      /* === IMPACT BOX (why this matters) === */
      .impact-box {
        background: linear-gradient(135deg, #1a5276, #1f618d);
        color: #ffffff;
        padding: 18px 26px;
        border-radius: 8px;
        font-size: 14.5px;
        line-height: 1.9;
        margin: 12px 0;
      }
      .impact-box strong {
        color: #aed6f1;
      }

      /* === ABOUT TAB CARDS === */
      .info-card {
        background: #ffffff;
        border: 1px solid #d5dbdb;
        border-left: 5px solid #1a5276;
        border-radius: 6px;
        padding: 16px 20px;
        margin-bottom: 16px;
        font-size: 14.5px;
        line-height: 1.85;
      }
      .info-card h4 {
        color: #1a5276;
        margin: 0 0 8px 0;
        font-size: 15px;
        font-weight: 700;
      }

      /* === VARIABLES TABLE === */
      .var-table {
        width: 100%;
        border-collapse: collapse;
        font-size: 14px;
        margin-top: 10px;
      }
      .var-table th {
        background: #1a5276;
        color: white;
        padding: 9px 13px;
        text-align: left;
      }
      .var-table td {
        padding: 8px 13px;
        border: 1px solid #dde;
      }
      .var-table tr:nth-child(even) td {
        background: #f8f9fa;
      }

      /* === CHART PANEL HEADERS === */
      .chart-title {
        font-size: 14px !important;
        font-weight: 600 !important;
        color: #1a5276 !important;
      }

      /* === MOBILE RESPONSIVE === */
      @media (max-width: 768px) {
        .value-box .value { font-size: 24px !important; }
        .value-box .caption { font-size: 11px !important; }
        .callout { font-size: 13px; }
        .impact-box { font-size: 13.5px; padding: 14px 18px; }
        .info-card { font-size: 13.5px; }
      }
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(DT)
library(htmltools)

# ── Colour palette ─────────────────────────────────────────────
gov_colors <- c(
  "Abyan"    = "#185FA5",
  "Aden"     = "#1D9E75",
  "Ad'Dhale" = "#D85A30",
  "Lahj"     = "#7F77DD"
)
year_colors <- c("2024" = "#2471A3", "2025" = "#C0392B")

month_order <- c("January","February","March","April","May","June",
                 "July","August","September","October","November","December")
month_abbr  <- c("Jan","Feb","Mar","Apr","May","Jun",
                 "Jul","Aug","Sep","Oct","Nov","Dec")

# ── Dataset 1: Monthly VPD cases ───────────────────────────────
monthly_gov <- tribble(
  ~Year,  ~Month,       ~Governorate, ~Cases,
  "2024","October",   "Abyan",  6,
  "2024","November",  "Abyan", 11,
  "2024","December",  "Abyan",  4,
  "2025","January",   "Abyan", 21,
  "2025","February",  "Abyan", 31,
  "2025","March",     "Abyan", 20,
  "2025","April",     "Abyan",  0,
  "2025","May",       "Abyan",  0,
  "2025","June",      "Abyan",  0,
  "2025","July",      "Abyan",  0,
  "2025","August",    "Abyan",  0,
  "2025","September", "Abyan",  0,
  "2024","April",     "Ad'Dhale",  2,
  "2024","May",       "Ad'Dhale", 12,
  "2024","June",      "Ad'Dhale", 17,
  "2024","July",      "Ad'Dhale", 19,
  "2024","August",    "Ad'Dhale", 39,
  "2024","September", "Ad'Dhale", 30,
  "2024","October",   "Ad'Dhale", 13,
  "2024","November",  "Ad'Dhale", 17,
  "2024","December",  "Ad'Dhale", 19,
  "2025","January",   "Ad'Dhale", 28,
  "2025","February",  "Ad'Dhale", 36,
  "2025","March",     "Ad'Dhale", 17,
  "2025","April",     "Ad'Dhale", 26,
  "2025","May",       "Ad'Dhale", 23,
  "2025","June",      "Ad'Dhale", 15,
  "2025","July",      "Ad'Dhale", 31,
  "2025","August",    "Ad'Dhale", 24,
  "2025","September", "Ad'Dhale", 22,
  "2024","October",   "Aden", 10,
  "2024","November",  "Aden", 27,
  "2024","December",  "Aden", 18,
  "2025","January",   "Aden",  8,
  "2025","February",  "Aden",  6,
  "2025","March",     "Aden", 26,
  "2025","April",     "Aden",  0,
  "2025","May",       "Aden",  0,
  "2025","June",      "Aden",  0,
  "2025","July",      "Aden",  0,
  "2025","August",    "Aden",  0,
  "2025","September", "Aden",  0,
  "2024","November",  "Lahj", 12,
  "2024","December",  "Lahj",  6,
  "2025","January",   "Lahj", 40,
  "2025","February",  "Lahj",  8,
  "2025","March",     "Lahj", 19,
  "2025","April",     "Lahj", 17,
  "2025","May",       "Lahj", 37,
  "2025","June",      "Lahj", 17,
  "2025","July",      "Lahj", 27,
  "2025","August",    "Lahj", 13,
  "2025","September", "Lahj", 11
) %>%
  mutate(
    Month      = factor(Month, levels = month_order, ordered = TRUE),
    Month_abbr = month_abbr[as.integer(Month)]
  )

# ── Dataset 2: Disease summary ──────────────────────────────────
summary_tbl <- tribble(
  ~Governorate, ~Disease,        ~`2024 Cases`, ~`2025 Cases`,
  "Ad'Dhale","Typhoid Fever",  125, 173,
  "Aden",    "Typhoid Fever",   32,  10,
  "Abyan",   "Typhoid Fever",   10,  13,
  "Lahj",    "Typhoid Fever",    4,  18,
  "Lahj",    "Chickenpox",       3,  56,
  "Abyan",   "Chickenpox",       4,  40,
  "Ad'Dhale","Chickenpox",      12,  10,
  "Aden",    "Chickenpox",      13,  13,
  "Lahj",    "Measles",          0,  32,
  "Ad'Dhale","Measles",          5,  15,
  "Abyan",   "Measles",          1,  10,
  "Aden",    "Measles",          1,   4,
  "Ad'Dhale","Mumps",           17,  16,
  "Aden",    "Mumps",            8,  10,
  "Abyan",   "Mumps",            6,   9,
  "Lahj",    "Mumps",            2,  21,
  "Ad'Dhale","Hepatitis A",      9,   8,
  "Lahj",    "Hepatitis A",      4,  24,
  "Abyan",   "Hepatitis A",      0,   2,
  "Aden",    "Hepatitis A",      0,   2,
  "Lahj",    "Pertussis",        2,  33,
  "Aden",    "Pertussis",        1,   0,
  "Lahj",    "Hepatitis B",      3,   4,
  "Aden",    "Rubella",          0,   1,
  "Lahj",    "Meningitis",       0,   1
) %>%
  mutate(
    Change     = `2025 Cases` - `2024 Cases`,
    `% Change` = ifelse(
      `2024 Cases` == 0, NA_real_,
      round((Change / `2024 Cases`) * 100, 1)
    )
  ) %>%
  arrange(desc(`2025 Cases`))

# ── Dataset 3: Measles doses vs cases ──────────────────────────
vacc_compare <- tribble(
  ~Governorate, ~Year, ~Measles_Doses, ~Measles_Cases,
  "Abyan",    "2024",  504,  7,
  "Abyan",    "2025",  842, 61,
  "Ad'Dhale", "2024",  480, 15,
  "Ad'Dhale", "2025", 1250, 45,
  "Aden",     "2024",   45,  3,
  "Aden",     "2025",   32, 11,
  "Lahj",     "2024",  225,  0,
  "Lahj",     "2025",  823, 32
)

# ── KPI values ──────────────────────────────────────────────────
total_2024 <- monthly_gov %>% filter(Year=="2024") %>% pull(Cases) %>% sum()
total_2025 <- monthly_gov %>% filter(Year=="2025") %>% pull(Cases) %>% sum()
all_govs   <- c("Abyan","Aden","Ad'Dhale","Lahj")
all_years  <- c("2025","2024")
```


<!-- ============================================================
     TAB 1: DASHBOARD
============================================================ -->

# Dashboard {data-icon="fa-chart-line"}

## Row {data-height=120}

### Total VPD Cases in 2024

```{r}
valueBox(total_2024, icon="fa-calendar-check", color="#2471A3",
  caption="Total VPD Cases — 2024 (available months only)")
```

### Total VPD Cases in 2025 (Jan–Sep)

```{r}
valueBox(total_2025, icon="fa-virus", color="#C0392B",
  caption="Total VPD Cases — 2025 (January to September)")
```

### Lahj: Case Increase 2024 to 2025

```{r}
valueBox("+950%", icon="fa-arrow-trend-up", color="#7F77DD",
  caption="Lahj: 18 cases in 2024 to 189 cases in 2025")
```

### Governorates Monitored

```{r}
valueBox(4, icon="fa-map-marker-alt", color="#1D9E75",
  caption="Abyan, Aden, Ad'Dhale, and Lahj")
```

## Row {data-height=90}

### {.no-padding}

<div class="callout">
<strong>How to read Widget 1A (left):</strong>
Each coloured line represents one governorate. Use the <strong>Year dropdown</strong> to switch between 2024 and 2025. Hover over any data point to see the governorate name, month, year, and exact case count.
&nbsp;&nbsp;|&nbsp;&nbsp;
<strong>Widget 1B (right):</strong>
Use the <strong>Governorate dropdown</strong> to select one governorate. The solid red line shows 2025 cases and the dashed blue line shows 2024 cases month by month. Hover for full detail.
</div>

## Row {data-height=520}

### Widget 1A: Monthly VPD Trends by Governorate — Select a Year {data-width=6}

```{r widget1a}
fig1a   <- plot_ly()
tmeta_a <- list()
tidx_a  <- 0

for (yr in all_years) {
  for (gov in all_govs) {
    sub <- monthly_gov %>%
      filter(Governorate == gov, Year == yr) %>%
      arrange(Month)

    tidx_a <- tidx_a + 1
    tmeta_a[[tidx_a]] <- yr

    fig1a <- fig1a %>%
      add_trace(
        data        = sub,
        x           = ~Month_abbr,
        y           = ~Cases,
        type        = "scatter",
        mode        = "lines+markers",
        name        = gov,
        legendgroup = gov,
        showlegend  = (yr == "2025"),
        visible     = (yr == "2025"),
        line   = list(color = gov_colors[[gov]], width = 3),
        marker = list(color = gov_colors[[gov]], size = 10,
                      symbol = "circle",
                      line = list(color = "white", width = 2)),
        hovertemplate = paste0(
          "<b>", gov, "</b><br>",
          "Month: <b>%{x}</b><br>",
          "Year: <b>", yr, "</b><br>",
          "New VPD Cases: <b>%{y}</b>",
          "<extra></extra>"
        )
      )
  }
}

vis_yr <- function(y) sapply(tmeta_a, function(x) x == y)

fig1a %>%
  layout(
    title = list(
      text = "<b>Monthly New VPD Cases by Governorate</b><br><sup>Yemen, CARE-Supported Governorates | Use the dropdown to select a year</sup>",
      font = list(size = 15, family = "Arial, sans-serif", color = "#1a2f45"),
      x = 0.5
    ),
    updatemenus = list(list(
      type = "dropdown", direction = "down",
      x = 0.01, xanchor = "left",
      y = 1.10, yanchor = "top",
      bgcolor = "white", bordercolor = "#2471A3", borderwidth = 2,
      font = list(size = 13),
      buttons = list(
        list(method = "restyle",
             args   = list("visible", as.list(vis_yr("2025"))),
             label  = "  Year: 2025  "),
        list(method = "restyle",
             args   = list("visible", as.list(vis_yr("2024"))),
             label  = "  Year: 2024  ")
      )
    )),
    xaxis = list(
      title = list(text = "Month",
                   font = list(size = 14, color = "#2c3e50")),
      tickfont = list(size = 12, color = "#2c3e50"),
      categoryorder = "array",
      categoryarray = month_abbr,
      showgrid = TRUE, gridcolor = "#eeeeee"
    ),
    yaxis = list(
      title = list(text = "New VPD Cases (count)",
                   font = list(size = 14, color = "#2c3e50")),
      tickfont = list(size = 12, color = "#2c3e50"),
      rangemode = "tozero",
      showgrid = TRUE, gridcolor = "#eeeeee"
    ),
    legend = list(
      title = list(text = "<b>Governorate</b>", font = list(size = 13)),
      font  = list(size = 12),
      bgcolor = "rgba(255,255,255,0.95)",
      bordercolor = "#cccccc", borderwidth = 1
    ),
    hovermode  = "x unified",
    hoverlabel = list(bgcolor = "white", bordercolor = "#cccccc",
                      font = list(size = 13, color = "#2c3e50")),
    plot_bgcolor  = "#fafafa",
    paper_bgcolor = "white",
    margin = list(t = 120, b = 50, l = 70, r = 20)
  ) %>%
  config(displayModeBar = TRUE,
         modeBarButtonsToRemove = c("lasso2d","select2d"),
         toImageButtonOptions = list(filename="vpd_trends", format="png"))
```

### Widget 1B: 2024 vs 2025 Monthly Cases — Select a Governorate {data-width=6}

```{r widget1b}
fig1b   <- plot_ly()
tmeta_b <- list()
tidx_b  <- 0

for (gov in all_govs) {
  for (yr in all_years) {
    sub <- monthly_gov %>%
      filter(Governorate == gov, Year == yr) %>%
      arrange(Month)

    tidx_b <- tidx_b + 1
    tmeta_b[[tidx_b]] <- gov

    fig1b <- fig1b %>%
      add_trace(
        data        = sub,
        x           = ~Month_abbr,
        y           = ~Cases,
        type        = "scatter",
        mode        = "lines+markers",
        name        = paste("Year", yr),
        legendgroup = yr,
        showlegend  = (gov == "Ad'Dhale"),
        visible     = (gov == "Ad'Dhale"),
        line = list(
          color = year_colors[[yr]],
          width = 3,
          dash  = ifelse(yr == "2025", "solid", "dash")
        ),
        marker = list(
          color  = year_colors[[yr]],
          size   = 10,
          symbol = ifelse(yr == "2025", "circle", "diamond"),
          line   = list(color = "white", width = 2)
        ),
        hovertemplate = paste0(
          "<b>", gov, " \u2014 ", yr, "</b><br>",
          "Month: <b>%{x}</b><br>",
          "New VPD Cases: <b>%{y}</b>",
          "<extra></extra>"
        )
      )
  }
}

vis_gov  <- function(g) sapply(tmeta_b, function(x) x == g)
gov_btns <- lapply(all_govs, function(g) {
  list(method = "restyle",
       args   = list("visible", as.list(vis_gov(g))),
       label  = paste0("  ", g, "  "))
})

fig1b %>%
  layout(
    title = list(
      text = "<b>Monthly VPD Cases: 2024 vs 2025 by Governorate</b><br><sup>Select a governorate | Solid line = 2025 &nbsp; Dashed line = 2024</sup>",
      font = list(size = 15, family = "Arial, sans-serif", color = "#1a2f45"),
      x = 0.5
    ),
    updatemenus = list(list(
      type = "dropdown", direction = "down",
      x = 0.01, xanchor = "left",
      y = 1.10, yanchor = "top",
      bgcolor = "white", bordercolor = "#C0392B", borderwidth = 2,
      font = list(size = 13),
      buttons = gov_btns
    )),
    xaxis = list(
      title = list(text = "Month",
                   font = list(size = 14, color = "#2c3e50")),
      tickfont = list(size = 12, color = "#2c3e50"),
      categoryorder = "array",
      categoryarray = month_abbr,
      showgrid = TRUE, gridcolor = "#eeeeee"
    ),
    yaxis = list(
      title = list(text = "New VPD Cases (count)",
                   font = list(size = 14, color = "#2c3e50")),
      tickfont = list(size = 12, color = "#2c3e50"),
      rangemode = "tozero",
      showgrid = TRUE, gridcolor = "#eeeeee"
    ),
    legend = list(
      title = list(text = "<b>Year</b>", font = list(size = 13)),
      font  = list(size = 12),
      bgcolor = "rgba(255,255,255,0.95)",
      bordercolor = "#cccccc", borderwidth = 1
    ),
    hovermode  = "x unified",
    hoverlabel = list(bgcolor = "white", bordercolor = "#cccccc",
                      font = list(size = 13, color = "#2c3e50")),
    plot_bgcolor  = "#fafafa",
    paper_bgcolor = "white",
    margin = list(t = 120, b = 50, l = 70, r = 20)
  ) %>%
  config(displayModeBar = TRUE,
         modeBarButtonsToRemove = c("lasso2d","select2d"),
         toImageButtonOptions = list(filename="vpd_comparison", format="png"))
```

## Row {data-height=100}

### {.no-padding}

<div class="impact-box">
<strong>Why this matters:</strong>
Vaccine-preventable diseases are surging across CARE-supported governorates in Yemen, with Lahj showing a 950% increase in VPD cases between 2024 and 2025. This sharp rise is directly connected to disruptions in routine immunization caused by ongoing conflict.
<strong>Strengthening EPI outreach and restoring cold-chain infrastructure in these districts is a critical, measurable, and cost-effective priority for humanitarian health actors working in Yemen.</strong>
</div>


<!-- ============================================================
     TAB 2: DISEASE TABLE & VACCINATION
============================================================ -->

# Disease Table & Vaccination {data-icon="fa-table"}

## Row {data-height=90}

### {.no-padding}

<div class="callout">
<strong>Widget 2 (left):</strong>
This table shows new VPD case counts by disease and governorate for 2024 and 2025. Use the <strong>search boxes at the top of each column</strong> to filter, and click any column header to sort.
<span style="color:#A32D2D;font-weight:bold;">Red</span> values mean cases increased from 2024 to 2025;
<span style="color:#1D9E75;font-weight:bold;">green</span> means they decreased.
Hover over any cell to see the column label.
&nbsp;&nbsp;|&nbsp;&nbsp;
<strong>Widget 3 (right):</strong>
The solid bars show measles vaccination doses given. The faded bars show measles cases reported. Blue = 2024, Red = 2025. Hover over any bar for full detail.
</div>

## Row {data-height=520}

### Widget 2: VPD Cases by Disease and Governorate (2024 vs 2025) {data-width=6}

```{r widget2}
dt_display <- summary_tbl %>%
  mutate(
    `% Change` = case_when(
      is.na(`% Change`) & `2024 Cases` == 0 & `2025 Cases` == 0 ~ "\u2014",
      is.na(`% Change`) & `2024 Cases` == 0 ~ "New in 2025",
      TRUE ~ paste0(`% Change`, "%")
    )
  )

datatable(
  dt_display,
  caption = tags$caption(
    style = "caption-side:top; text-align:left; font-size:15px;
             font-weight:bold; color:#1a5276; padding-bottom:10px;",
    "New VPD Cases by Governorate and Disease \u2014 Yemen 2024 vs 2025"
  ),
  filter   = "top",
  rownames = FALSE,
  options  = list(
    dom        = "frtip",
    pageLength = 25,
    scrollX    = TRUE,
    columnDefs = list(
      list(width = "120px", targets = 0),
      list(width = "140px", targets = 1),
      list(className = "dt-right", targets = c(2,3,4,5)),
      list(width = "88px", targets = c(2,3,4,5))
    )
  ),
  callback = DT::JS("
    table.on('mouseover', 'td', function() {
      var cols = ['Governorate','Disease',
                  '2024 Cases','2025 Cases','Change','% Change'];
      var i = table.cell(this).index().column;
      $(this).attr('title', cols[i] + ': ' + $(this).text());
    });
  "),
  class = "cell-border stripe hover"
) %>%
  formatStyle("2025 Cases",
    background = styleColorBar(
      c(0, max(summary_tbl[["2025 Cases"]], na.rm=TRUE)+5), "#B5D4F4"),
    backgroundSize = "90% 65%",
    backgroundRepeat = "no-repeat",
    backgroundPosition = "center"
  ) %>%
  formatStyle("Change",
    color = styleInterval(c(-0.001, 0.001),
                          c("#1D9E75", "gray", "#A32D2D")),
    fontWeight = "bold"
  ) %>%
  formatStyle("% Change",
    fontWeight = "bold"
  ) %>%
  formatStyle(columns=0:5, fontSize="14px")
```

### Widget 3: Measles Vaccination Doses vs Measles Cases {data-width=6}

```{r widget3}
plot_ly() %>%
  add_bars(
    data = vacc_compare %>% filter(Year=="2024"),
    x = ~Governorate, y = ~Measles_Doses,
    name = "Doses Given \u2014 2024",
    marker = list(color="#2471A3", opacity=0.90,
                  line=list(color="white",width=1.5)),
    hovertemplate = paste0(
      "<b>Doses Given \u2014 2024</b><br>",
      "Governorate: <b>%{x}</b><br>",
      "Doses Administered: <b>%{y}</b>",
      "<extra></extra>")
  ) %>%
  add_bars(
    data = vacc_compare %>% filter(Year=="2025"),
    x = ~Governorate, y = ~Measles_Doses,
    name = "Doses Given \u2014 2025",
    marker = list(color="#C0392B", opacity=0.90,
                  line=list(color="white",width=1.5)),
    hovertemplate = paste0(
      "<b>Doses Given \u2014 2025</b><br>",
      "Governorate: <b>%{x}</b><br>",
      "Doses Administered: <b>%{y}</b>",
      "<extra></extra>")
  ) %>%
  add_bars(
    data = vacc_compare %>% filter(Year=="2024"),
    x = ~Governorate, y = ~Measles_Cases,
    name = "Cases Reported \u2014 2024",
    marker = list(color="#2471A3", opacity=0.35,
                  line=list(color="#2471A3",width=1.5)),
    hovertemplate = paste0(
      "<b>Cases Reported \u2014 2024</b><br>",
      "Governorate: <b>%{x}</b><br>",
      "Cases: <b>%{y}</b>",
      "<extra></extra>")
  ) %>%
  add_bars(
    data = vacc_compare %>% filter(Year=="2025"),
    x = ~Governorate, y = ~Measles_Cases,
    name = "Cases Reported \u2014 2025",
    marker = list(color="#C0392B", opacity=0.35,
                  line=list(color="#C0392B",width=1.5)),
    hovertemplate = paste0(
      "<b>Cases Reported \u2014 2025</b><br>",
      "Governorate: <b>%{x}</b><br>",
      "Cases: <b>%{y}</b>",
      "<extra></extra>")
  ) %>%
  layout(
    title = list(
      text = "<b>Measles Vaccination Doses vs Measles Cases Reported</b><br><sup>By Governorate, Yemen 2024 vs 2025 | Solid bars = doses given; faded bars = cases reported</sup>",
      font = list(size = 14, family = "Arial, sans-serif", color = "#1a2f45"),
      x = 0.5
    ),
    barmode = "group",
    bargap  = 0.22,
    xaxis = list(
      title = list(text = "Governorate",
                   font = list(size = 14, color = "#2c3e50")),
      tickfont = list(size = 12, color = "#2c3e50"),
      showgrid = FALSE
    ),
    yaxis = list(
      title = list(text = "Count (doses or cases)",
                   font = list(size = 14, color = "#2c3e50")),
      tickfont = list(size = 12, color = "#2c3e50"),
      rangemode = "tozero",
      showgrid = TRUE, gridcolor = "#eeeeee"
    ),
    legend = list(
      font = list(size = 12),
      bgcolor = "rgba(255,255,255,0.95)",
      bordercolor = "#cccccc", borderwidth = 1,
      orientation = "h", x = 0, y = -0.28
    ),
    hovermode  = "closest",
    hoverlabel = list(bgcolor = "white", bordercolor = "#cccccc",
                      font = list(size = 13, color = "#2c3e50")),
    plot_bgcolor  = "#fafafa",
    paper_bgcolor = "white",
    margin = list(t = 90, b = 120, l = 70, r = 20)
  ) %>%
  config(displayModeBar = TRUE,
         modeBarButtonsToRemove = c("lasso2d","select2d"),
         toImageButtonOptions = list(filename="measles_doses_cases", format="png"))
```

## Row {data-height=100}

### {.no-padding}

<div class="impact-box">
<strong>Why this matters:</strong>
Vaccine-preventable diseases are surging across CARE-supported governorates in Yemen, with Lahj showing a 950% increase in VPD cases between 2024 and 2025. This sharp rise is directly connected to disruptions in routine immunization caused by ongoing conflict.
<strong>Strengthening EPI outreach and restoring cold-chain infrastructure in these districts is a critical, measurable, and cost-effective priority for humanitarian health actors working in Yemen.</strong>
</div>


<!-- ============================================================
     TAB 3: ABOUT THE DATA
     Rubric: separate tab, data source, sample size, collection
     method, study population, time period — max 5 sentences
============================================================ -->

# About the Data {data-icon="fa-info-circle"}

## Row

### Dataset Description {data-width=7}

```{r about, echo=FALSE}
div(
  style = "padding:26px; font-size:14.5px; line-height:1.9; color:#2c3e50;",

  h3(style="color:#1a5276; border-bottom:2px solid #2980b9;
            padding-bottom:9px; margin-top:0; font-size:20px;",
     "About This Dataset"),

  div(class="info-card",
    h4("Data Source"),
    p("This dashboard uses CARE Yemen's internal ",
      strong("OPD (Outpatient Department) surveillance records"),
      " and ", strong("EPI (Expanded Programme on Immunization) monthly data entry sheets;"),
      " the dataset is ", strong("confidential"),
      " and not publicly available, and was shared under a data-use agreement
      for academic research purposes only.")
  ),

  div(class="info-card",
    h4("Data Collection Method"),
    p("Data were collected through ", strong("passive OPD disease surveillance"),
      " where health facility staff recorded new VPD case counts on a monthly basis,
      alongside ", strong("active EPI tracking"),
      " of vaccination doses administered per governorate each month.")
  ),

  div(class="info-card",
    h4("Study Population"),
    p("Patients who presented to CARE-supported health facilities in four
      governorates of Yemen — Abyan, Aden, Ad'Dhale, and Lahj — with symptoms
      of vaccine-preventable diseases, including Typhoid Fever, Measles,
      Chickenpox, Mumps, Hepatitis A/B, Pertussis, Rubella, and Meningitis.")
  ),

  div(class="info-card",
    h4("Time Period and Location"),
    p("The data cover ", strong("April 2024 through September 2025"),
      " across four CARE-supported governorates in Yemen. Months with
      incomplete surveillance coverage — January to March 2024 and
      October to December 2025 — are excluded from trend analyses.")
  ),

  div(class="info-card",
    h4("Sample Size"),
    p("A total of ", strong("498 new VPD case records"),
      " across 25 governorate-disease combinations are included in the
      disease-level table. The monthly trend charts are based on 51
      governorate-month-year data points across the four governorates.")
  ),

  hr(style="border-top:1px solid #dee2e6; margin:22px 0;"),

  h4(style="color:#1a5276; font-size:16px;", "Research Question"),
  p(em("How were disruptions in routine immunization associated with changes
       in reported vaccine-preventable disease incidence in CARE-supported
       districts in Yemen between 2024 and 2025?")),

  h4(style="color:#1a5276; margin-top:20px; font-size:16px;", "Key Variables"),
  tags$table(class="var-table",
    tags$thead(tags$tr(
      tags$th("Variable"), tags$th("Type"), tags$th("Description")
    )),
    tags$tbody(
      tags$tr(tags$td(strong("New VPD Cases")), tags$td("Count"),
              tags$td("Monthly new VPD patients at CARE OPD facilities")),
      tags$tr(tags$td(strong("Governorate")), tags$td("Categorical"),
              tags$td("Abyan, Aden, Ad'Dhale, or Lahj")),
      tags$tr(tags$td(strong("Year")), tags$td("Categorical"),
              tags$td("2024 or 2025")),
      tags$tr(tags$td(strong("Month")), tags$td("Ordered categorical"),
              tags$td("January through December")),
      tags$tr(tags$td(strong("Disease")), tags$td("Categorical"),
              tags$td("Typhoid, Measles, Chickenpox, Mumps, Hep A/B, Pertussis, Rubella, Meningitis")),
      tags$tr(tags$td(strong("Measles Doses")), tags$td("Count"),
              tags$td("1st + 2nd measles dose administered — from EPI records")),
      tags$tr(tags$td(strong("% Change")), tags$td("Numeric (%)"),
              tags$td("Percentage change from 2024 to 2025, rounded to 1 decimal place"))
    )
  )
)
```

### Limitations & Author {data-width=3}

```{r limits, echo=FALSE}
div(
  style = "padding:24px; font-size:14px; line-height:1.9; color:#2c3e50;",

  h4(style="color:#922b21; margin-top:0; font-size:15px;",
     "\u26a0\ufe0f Important Limitations"),
  tags$ul(style="padding-left:18px; line-height:2.1;",
    tags$li("The surveillance data are ", strong("passive"),
      " which means true case counts are likely underreported, especially
      in areas with restricted access due to the conflict."),
    tags$li("Zero cases for Abyan and Aden in April–September 2025 may reflect ",
      strong("data collection gaps"), " rather than an actual absence of disease."),
    tags$li("The relationship between vaccination doses and cases in Widget 3 is ",
      strong("descriptive, not causal"), " — displacement, immunity gaps, and
      access barriers all independently affect disease incidence."),
    tags$li("Population denominators are not available, so it is not possible to
      calculate ", strong("incidence rates"), " from this dataset."),
    tags$li("Surveillance coverage start dates differ by governorate, which
      limits direct cross-governorate comparisons for 2024.")
  ),

  hr(style="border-top:1px solid #dee2e6; margin:16px 0;"),

  h4(style="color:#1a5276; font-size:15px;", "\U0001f4e6 R Packages"),
  tags$ul(style="padding-left:18px; line-height:2.2;",
    tags$li(strong("flexdashboard"), " — dashboard layout and tab structure"),
    tags$li(strong("plotly"), " — interactive charts with dropdown menus and tooltips"),
    tags$li(strong("DT"), " — interactive searchable and sortable table"),
    tags$li(strong("tidyverse"), " — data wrangling and transformation"),
    tags$li(strong("htmltools"), " — custom HTML in the About tab")
  ),

  hr(style="border-top:1px solid #dee2e6; margin:16px 0;"),

  h4(style="color:#1a5276; font-size:15px;", "\U0001f9d1\u200d\U0001f4bb Author"),
  p(strong("Akanshya Dash"), br(),
    "DATA 555 — Data Science Capstone", br(),
    "Rollins School of Public Health", br(),
    "Emory University | Spring 2026")
)
```
